---
title: "Nickel LC50 Experiment 2"
author: "Celeste Valdivia"
date: "2024-04-19"
categories: [study plans, nickel chloride]
---
# Assigning Animals Randomly to different treatments

```{r}
set.seed(3349)

conditions <- c("Control R1", 
                "Control R2", 
                "0.1 mg/L R1", 
                "0.1 mg/L R2", 
                "1 mg/L R1", 
                "1 mg/L R2", 
                "10 mg/L R1", 
                "10 mg/L R2", 
                "100 mg/L R1", 
                "100 mg/L R2", 
                "1000 mg/L R1", 
                "1000 mg/L R2")
animal_id <- c("DM022024_C01", 
               "DM022024_C02", 
               "DM022024_C03", 
               "DM022024_C04*", 
               "DM022024_C05", 
               "DM022024_C06", 
               "DM022024_C07", 
               "DM022024_C08",
               "DM022024_C09",
               "DM022024_C10", 
               "DM022024_C11", 
               "DM022024_C12")

treatment_assignments <- data.frame(
  Conditions = sample(conditions),
  animal_ID = sample(animal_id)
)
print(treatment_assignments)

#save assignments in csv for downstream use as metadata
```

```{r}
setwd('..')
write.csv(treatment_assignments, file = "treatment_assignments_LC50_E2.csv")
```

# Nickel Spiking Calculations
## Nickel Chloride Dilutions

Review resources prior to handling nickel chloride:

-   Nickel chloride [SDS](https://www.fishersci.com/store/msds?partNumber=AC270510050&productDescription=NICKEL%28II%29+CHLORIDE+HEXA+5KG&vendorId=VN00032119&countryCode=US&language=en).

-   [Chemical resistance of gloves.](https://amo-csd.lbl.gov/downloads/Chemical%20Resistance%20of%20Gloves.pdf)

All handling of stock solution and treatments to be done in a chemical fume hood. Once spiked water has been distributed, add animals, parafilm jars, place in secondary container, and move to environmental chamber.

### Required Volume of Artificial Seawater

Calculations for spiking artificial seawater. 225 mL of seawater required to mostly submerge glass slide in wide mouth 16 oz Kerr mason jar.

Considerations:

-   2 replicates

-   225 mL per jar

-   6 conditions

-   Account for 10% loss of water during transferring

-   Additional 15 mL required for analytical chemistry

$$(2 * 225 * 6) + (15*6) + (0.1*(2*225*6)) + (0.1*(15*6)) = 3,069$$

Need 3,069 mL of artificial seawater total for exposure, particularly, 511.5 mL of artificial seawater per condition.

### Calculations

Stock solution of NiCl~2~ (10 g/L) made 04/08/24.

-   NiCl~2~ (5g)

-   Suspended in MilliQ water (500 mL)

#### 0.1 mg/L

$$ C_1*V_1 = C_2*V_2$$ $$ 10 g/L * V_1 = 0.0001g/L*0.5115L$$ $$V_1 = 0.000005115 L = 5.115 uL$$ 
Spike 511.5 mL of aerated, artificial seawater (18 C) with 5.115 uL of stock NiCl~2~ solution. Change in salinity negligible (+/- 1 PPT).

#### 1 mg/L

$$ C_1*V_1 = C_2*V_2$$ $$ 10 g/L * V_1 = 0.001g/L*0.5115L$$ $$V_1 = 0.00005115 L = 51.15 uL$$ 

Spike 511.5 mL of aerated, artificial seawater (18 C) with 51.15 uL of stock NiCl~2~ solution. Change in salinity negligible (+/- 1 PPT).

#### 10 mg/L

$$ C_1*V_1 = C_2*V_2$$ $$ 10 g/L * V_1 = 0.01g/L*0.5115L$$ $$V_1 = 0.0005115 L = 511.5 uL$$ 

Spike 511.5 mL of aerated, artificial seawater (18 C) with 511.5 uL of stock NiCl~2~ solution. Change in salinity negligible (+/- 1 PPT).

#### 100 mg/L

$$ C_1*V_1 = C_2*V_2$$ $$ 10 g/L * V_1 = 0.1g/L*0.5115L$$ $$V_1 = 0.005115 L = 5.115 mL$$ 

Spike 511.5 mL of aerated, artificial seawater (18 C) with 5.115 mL of stock NiCl~2~ solution. Change in salinity negligible (+/- 1 PPT).

#### 1000 mg/L

$$ C_1*V_1 = C_2*V_2$$ $$ 10 g/L * V_1 = 1g/L*0.5115L$$ $$V_1 = 0.0759 L = 51.15 mL$$ 

Spike 511.5 mL of aerated, artificial seawater (18 C) with 51.15 mL of stock NiCl~2~ solution. Change in salinity negligible (+/- 1 PPT).
