---
title: "Exposure 2 *in vivo* NiCl~2~ x *Botryllus schlosseri*"
author: "Celeste Valdivia"
date: "2023-11-13"
categories: [study plans, nickel chloride]
image: ""
---

```{r setup, include=TRUE, warning=FALSE}
library(knitr)
library(tidyverse)
library(tidyr)
library(dplyr)
library(hrbrthemes)
library(ggplot2)
knitr::opts_chunk$set(echo = TRUE)
```

# Experimental Planning

## Animal Handling Prior to Exposure

Divided animals into separate system, 3 days prior to start of exposure, 8 days acclimated to the recirculating system.

Only 9 individuals had properly adhered systems. Will reduce the number of replicates.

## Randomize Treatment 

Here we will have R randomly assign each animal to a treatment. 

```{r}
set.seed(3789)

conditions <- c("Control R1", "Control R2", "Control R3", "5 mg/L R1", "5 mg/L R2", "5 mg/L R3", "45 mg/L R1", "45 mg/L R2", "45 mg/L R3")
animal_id <- c("S266C001", "S259C001", "S271C001", "S263C002", "S251C001", "S272C001", "S256C001", "S270C001", "S267C001")
treatment_assignments_exp2 <- data.frame(
  Conditions = sample(conditions),
  animal_ID = sample(animal_id)
)
print(treatment_assignments_exp2)

#save assignments in csv for downstream use as metadata
write.csv(treatment_assignments_exp2, file = "treatment_assignments.csv")
```

## Nickel Chloride Dilutions

Review resources prior to handling nickel chloride:

-   Nickel chloride [SDS](https://www.fishersci.com/store/msds?partNumber=AC270510050&productDescription=NICKEL%28II%29+CHLORIDE+HEXA+5KG&vendorId=VN00032119&countryCode=US&language=en).

-   [Chemical resistance of gloves.](https://amo-csd.lbl.gov/downloads/Chemical%20Resistance%20of%20Gloves.pdf)

All handling of stock solution and treatments to be done in a chemical fume hood. Once spiked water has been distributed, add animals, parafilm jars, place in secondary container, and move to environmental chamber.

### Required Volume of Artificial Seawater

Calculations for spiking artificial seawater. 225 mL of seawater required to mostly submerge glass slide in wide mouth 16 oz Kerr mason jar.

Considerations:

-   3 replicates

-   225 mL per jar

-   3 conditions

-   Account for 10% loss of water during transferring

$$(3* 220 * 3) + (0.1*(3*220*3)) = 2,178$$

Need 2,178 mL of artificial seawater total for exposure, particularly, 726 mL of artificial seawater per condition.

### Calculations

Stock solution of NiCl~2~ made by undergraduate Gina Jones on 02/08/23.

-   NiCl~2~

-   Suspended in MilliQ water (0 PPT)

#### Low Dose: 5 mg/L

$$ C_1*V_1 = C_2*V_2$$ $$ 1 g/L * V_1 = 0.005g/L*0.726L$$ $$V_1 = 0.00363 L = 3.63 mL$$ Spike 726 mL of aerated, artificial seawater (16 C) with 4.95 mL of stock NiCl~2~ solution. Change in salinity negligible (+/- 1 PPT).

#### High Dose: 45 mg/L

$$ C_1*V_1 = C_2*V_2$$ $$ 1 g/L * V_1 = 0.045g/L*0.726L$$ $$V_1 = 0.03267 L = 32.67 mL$$ Spike 726 mL of aerated, artificial seawater (16 C) with 44.55 mL of stock NiCl~2~ solution. Change in salinity negligible (+/- 1 PPT).



# Protocol

1.  Subdivide animals 3 days prior to exposure start.

2.  Make sufficient volume of artificial seawater at appropriate salinity, 25 PPT, to match recirculating seawater system parameters. In this case for a 3 x 4 exposure at 250 mL, will require about 7 L of artificial seawater.

-   30.8 g / L of Red Sea Salt

3.  Rinse all glassware to be used during exposure as follows:

-   Triplicate rinse with lab detergent

-   Triplicate rinse with tap water

-   Triplicate rinse with DI water

4.  Add lab bench paper to environmental chambers top rack to reduce light intensity. Include 4 slits in bench paper to allow for air flow.

5.  Add 1000 mL of artificial seawater to three 1 L beakers and place in environmental chamber (set to ambient 23 C to achieve 16 C for water) with gentle aeration (targeting \~ 67 % DO) in preparation the night before exposure.

#### *Day of Exposure Commencement*

6.  Take 10 mL from each 1 L beaker and document temperature, DO, pH, salinity, and ammonia of artificial seawater prior to spiking the water.

7.  In a chemical fume hood add the appropriate volume to achieve desired concentration of nickel chloride concentration to each of the 1 L beakers using a 25 mL or 50 mL serological pipette. Be sure to homogenize solution well with the serological pipette. Distribute 225 mL of each stock to the appropriate mason jar or beaker using designated serological pipette.

8.  Transfer animals to a portable container with artificial seawater from the recirculating system. Image the animals using the Excelis stereomicroscope. Use a razor blade to clean around the animal and to clean glass slide of excess algae and debris.

-   Document stage, zooid number, cumulative system number.

9.  Distribute animals to appropriate jar in fume hood.

10. Parafilm jar and place in secondary container.

11. Transfer secondary container with jars to second rack of environmental chamber.

12. Leave animals for 24 hours.

#### *Takedown*

The following steps should be conducted in a chemical fume hood and iterated through one beaker at a time to avoid desiccation stress:

13. Rinse the entirety of the slide and animals with artificial seawter thorougly in quadruplicates to knock off any excess nickel chloride on the animal before transfering to save the tissue.

14. Using a stainless steel razor blade, remove one of the two cumulative systems on the glass slide. Use forceps to transfer cumulative system to appropriately labeled *piereced* snap cap 1.5 mL eppendorf PCR safe tube and snap freeze in liquid nitrogen for gene expression.

15. Using the same stainless steel razor blade, remove the other cumulative system from the slide and transfer it to tube for SOD assay. Snap freeze

16. Transfer samples containing cumulative system samples to -80 C freezer for long term storage.
